Meta RL analyses

### 2019-03-06 meta_RL_analyses.Rmd
Better baselining

result_subdirs = c(# "nometa",# "nometa_su_heldout", "nometa_random_holdout_longer", "nometa_only_one",  "nometa_only_one_ho",
                   "basic", #"basic_su_heldout", 
                   "basic_random_holdout_longer", # "basic_only_one", "basic_only_one_ho",
                   #"language", "language_simplified",
                   "language_noncompositional_old", "language_simplified_random_holdout")

@@

summary_plot = function(data,  x_var, y_var, color_var, baseline_data=NULL, palette="Set2") {
  p = ggplot(data, aes_string(x=x_var, y=y_var, color=color_var)) +
    geom_hline(yintercept=0., linetype=2, alpha=0.5) +
    geom_point(stat="summary", fun.y=mean, size=2) +
    geom_errorbar(stat="summary", 
                  fun.ymin=function(x) {mean(x) - 1.96*sd(x)/sqrt(length(x))},
                  fun.ymax=function(x) {mean(x) + 1.96*sd(x)/sqrt(length(x))},
                  width=0.25) +
    scale_color_brewer(palette=palette) +
    labs(x=name_var(x_var), y=name_var(y_var)) +
    guides(color=guide_legend(title=NULL))

  if (!is.null(baseline_data)) {
    p = p + 
      annotate("segment",
               x=(1:5)-0.5, xend=(1:5)+0.5,
               y=simple_baselines$expected_reward,
               yend=simple_baselines$expected_reward,
               linetype=1, alpha=0.5) 
  }

@@

```{r}
baseline_data = read_csv("baseline_data.csv") %>%
  mutate(game_type = str_extract(game, "high_card|match|straight_flush|pairs_and_high|sum_under"),
         losers = grepl("l_1", game),
         black_valuable = grepl("bv_1", game),
         suits_rule = grepl("sr_1", game))
```

@@

reward_d = reward_d %>%
  mutate(is_su_ho=grepl("su_heldout", run_type),
         is_only_one=grepl("only_one$", run_type),
         is_only_one_ho=grepl("only_one_ho", run_type),
         is_random_ho=grepl("random", run_type),
         is_default=!(is_su_ho | is_random_ho | is_only_one | is_only_one_ho)) %>%
  gather(game_str, reward, starts_with("game")) %>%
  mutate(game_type = str_extract(game_str, "high_card|match|straight_flush|pairs_and_high|sum_under"),
         losers = grepl("l_1", game_str),
         black_valuable = grepl("bv_1", game_str),
         suits_rule = grepl("sr_1", game_str),
         is_new_game = (is_default & game_str %in% new_game_strs) | (is_su_ho & game_str %in% new_game_strs_su_heldout) | (is_random_ho & game_str %in% new_game_strs_random_heldout) | (is_only_one & game_str %in% new_game_strs_only_one) | (is_only_one_ho & game_str %in% new_game_strs_only_one_ho),
         is_new_game = factor(is_new_game, labels=c("Trained", "Held out")))
```

@@

```{r}
simple_baselines = baseline_data %>% 
  filter(policy == "optimal") %>%
  group_by(game_type) %>%
  summarize(expected_reward = mean(expected_reward)) %>%
  ungroup()

# summary_plot(reward_d %>%
#                filter(epoch == 0),
#              "game_type", "reward", "is_new_game",
#              simple_baselines) +
#   facet_wrap(. ~ named_run_type)
```

```{r}
summary_plot(reward_d %>%
               filter(epoch == 0,
                      run_type == "basic"),
             "game_type", "reward", "is_new_game",
             simple_baselines) 
ggsave("~/Documents/grad/res/psych/ARRR/figures/meta_RL/basic.png", width=5, height=3)
```

```{r}
summary_plot(reward_d %>%
               filter(epoch == 0,
                      run_type %in% c("basic", "basic_random_holdout_longer")),
             "game_type", "reward", "is_new_game",
             simple_baselines) +
  facet_wrap(. ~ named_run_type) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("writing/figures/basic_meta_learning.png", width=5, height=3.5)
#ggsave("~/Documents/grad/res/psych/ARRR/figures/meta_RL/basic_random_holdout_half.png", width=5, height=3.5)
```

@@

  mutate(is_su_ho=grepl("su_heldout", run_type),
         is_only_one=grepl("only_one$", run_type),
         is_only_one_ho=grepl("only_one_ho", run_type),
         is_random_ho=grepl("random", run_type),
         is_default=!(is_su_ho | is_random_ho | is_only_one | is_only_one_ho)) %>%
  gather(game_str, reward, starts_with("game")) %>%
  mutate(game_type = str_extract(game_str, "high_card|match|straight_flush|pairs_and_high|sum_under"),
         losers = grepl("l_1", game_str),
         black_valuable = grepl("bv_1", game_str),
         suits_rule = grepl("sr_1", game_str),
         is_new_game = (is_default & game_str %in% new_game_strs) | (is_su_ho & game_str %in% new_game_strs_su_heldout) | (is_random_ho & game_str %in% new_game_strs_random_heldout) | (is_only_one & game_str %in% new_game_strs_only_one) | (is_only_one_ho & game_str %in% new_game_strs_only_one_ho))
```

@@

ggsave("~/Documents/grad/res/psych/ARRR/figures/meta_RL/basic_sweeping.png", width=5, height=3)
```

@@

```{r}
language_d = load_d(results_dir, c("language_noncompositional_old", 
                                   #"language", "language_simplified", "language_meta_nontoggling", "language_nontoggling_random_holdout", 
                                   "language_simplified_random_holdout"), num_runs, "new_language_rewards")

language_d = language_d %>%
  mutate(is_su_ho=grepl("su_heldout", run_type),
         is_only_one=grepl("only_one$", run_type),
         is_only_one_ho=grepl("only_one_ho", run_type),
         is_random_ho=grepl("random", run_type),
         is_default=!(is_su_ho | is_random_ho | is_only_one | is_only_one_ho)) %>%
  gather(game_str, reward, starts_with("game")) %>%
  mutate(game_type = str_extract(game_str, "high_card|match|straight_flush|pairs_and_high|sum_under"),
         losers = grepl("l_1", game_str),
         black_valuable = grepl("bv_1", game_str),
         suits_rule = grepl("sr_1", game_str),
         is_new_game = (is_default & game_str %in% new_game_strs) | (is_su_ho & game_str %in% new_game_strs_su_heldout) | (is_random_ho & game_str %in% new_game_strs_random_heldout) | (is_only_one & game_str %in% new_game_strs_only_one) | (is_only_one_ho & game_str %in% new_game_strs_only_one_ho))
```

```{r}
ggplot(language_d %>%
         filter(epoch == 0,
                is_new_game),
       aes(x=named_run_type, y=reward, fill=as.factor(run))) +
  geom_hline(yintercept=0.,
             linetype=2,
             alpha=0.5) +
  geom_violin(position="dodge") +
  geom_point(aes(group=as.factor(run)),
             size=3,
             stat="summary", fun.y="mean",
             position=position_dodge(width=0.9)) +
  scale_color_brewer(palette="Set2") + 
  ylim(-0.6, 0.6) +
  labs(x="Run type",
       y="Reward density") + 
  guides(fill=guide_legend(title="Run"))
ggsave("writing/figures/language_basic_by_run.png", width=5, height=3.5)
```

```{r}
summary_plot(language_d %>%
               filter(epoch == 0),
             "game_type", "reward", "is_new_game",
             simple_baselines) +
  facet_wrap(. ~ named_run_type) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("writing/figures/language_basic.png", width=5, height=3.5)
#ggsave("~/Documents/grad/res/psych/ARRR/figures/meta_RL/crappy_language.png", width=5, height=3)
```

@@

```{r}
meta_d = load_d(results_dir, c("basic", 
                               #"basic_su_heldout",
                               "basic_random_holdout_longer",# "basic_only_one", "basic_only_one_ho",
                               #"language", "language_drop_03", "language_drop_05", "language_noncompositional"
                               # "language", "language_simplified",
                               # "language_meta_nontoggling", "language_nontoggling_random_holdout", 
                               "language_noncompositional_old", "language_simplified_random_holdout"
                               ), num_runs, "new_meta_true_losses")

meta_d = meta_d %>%
  filter(epoch == 0) %>%
  mutate(is_su_ho=grepl("su_heldout", run_type),
         is_only_one=grepl("only_one$", run_type),
         is_only_one_ho=grepl("only_one_ho", run_type),
         is_random_ho=grepl("random", run_type),
         is_default=!(is_su_ho | is_random_ho | is_only_one)) %>%
  gather(map_str, reward, -c(run, run_type, named_run_type, epoch), -starts_with("is_")) %>%
  separate(map_str, c("meta_task", "source", "target"), sep="[\\.]+") %>%
  mutate(is_new = (is_default & (source %in% new_game_strs | target %in% new_game_strs)) | (is_random_ho & (source %in% new_game_strs_random_heldout | target %in% new_game_strs_random_heldout)) | (is_only_one & (source %in% new_game_strs_only_one | target %in% new_game_strs_only_one)) | (is_su_ho & (source %in% new_game_strs_su_heldout | target %in% new_game_strs_su_heldout))| (is_only_one_ho & (source %in% new_game_strs_only_one_ho | target %in% new_game_strs_only_one_ho)),
         is_new = factor(is_new, labels=c("Trained", "Held out")))
```

@@

ggsave("~/Documents/grad/res/psych/ARRR/figures/meta_RL/basic_basic_random_meta.png", width=5, height=3.5)
ggsave("writing/figures/meta_mapping.png", width=5, height=3.5)
```

@@

```{r}
lang_meta_d = load_d(results_dir, c(
                              # "language", "language_simplified", "language_meta_nontoggling", "language_nontoggling_random_holdout", 
                               "language_noncompositional_old", "language_simplified", "language_simplified_random_holdout"
                               ), num_runs, "new_language_meta_true_losses")

lang_meta_d = lang_meta_d %>%
  filter(epoch == 0) %>%
  mutate(is_su_ho=grepl("su_heldout", run_type),
         is_only_one=grepl("only_one$", run_type),
         is_only_one_ho=grepl("only_one_ho", run_type),
         is_random_ho=grepl("random", run_type),
         is_default=!(is_su_ho | is_random_ho | is_only_one)) %>%
  gather(map_str, reward, -c(run, run_type, named_run_type, epoch), -starts_with("is_")) %>%
  separate(map_str, c("meta_task", "source", "target"), sep="[\\.]+") %>%
  mutate(is_new = (is_default & (source %in% new_game_strs | target %in% new_game_strs)) | (is_random_ho & (source %in% new_game_strs_random_heldout | target %in% new_game_strs_random_heldout)) | (is_only_one & (source %in% new_game_strs_only_one | target %in% new_game_strs_only_one)) | (is_su_ho & (source %in% new_game_strs_su_heldout | target %in% new_game_strs_su_heldout))| (is_only_one_ho & (source %in% new_game_strs_only_one_ho | target %in% new_game_strs_only_one_ho)),
         is_new = factor(is_new, labels=c("Trained", "Held out")))
```

```{r}
summary_plot(lang_meta_d %>%
               filter(epoch == 0),
             "meta_task", "reward", "is_new") +
  facet_wrap(. ~ named_run_type)  +
  guides(color=guide_legend(title=NULL)) +
  labs(x="Game type", y="Average reward") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 ggsave("writing/figures/language_meta_mapping.png", width=5, height=3.5)
```
# embeddings

>>>>

result_subdirs = c(# "nometa",# "nometa_su_heldout", "nometa_random_holdout_longer", "nometa_only_one",  "nometa_only_one_ho",
                   "basic", #"basic_su_heldout", 
                   "basic_random_holdout_longer", # "basic_only_one", "basic_only_one_ho",
                   #"language", "language_simplified",
                   "language_noncompositional", "language_simplified", "language_meta_nontoggling",
                   "language_noncompositional_old", "language_simplified_random_holdout")

@@

summary_plot = function(data,  x_var, y_var, color_var, optimal_baseline_data=NULL, other_baseline_data=NULL, palette="Set2") {
  p = ggplot(data, aes_string(x=x_var, y=y_var, color=color_var)) +
    geom_hline(yintercept=0., linetype=2, alpha=0.5) +
    geom_point(stat="summary", fun.y=mean, size=2) +
    geom_errorbar(stat="summary", 
                  fun.ymin=function(x) {mean(x) - 1.96*sd(x)/sqrt(length(x))},
                  fun.ymax=function(x) {mean(x) + 1.96*sd(x)/sqrt(length(x))},
                  width=0.25) +
    scale_color_brewer(palette=palette) +
    labs(x=name_var(x_var), y=name_var(y_var)) +
    guides(color=guide_legend(title=NULL))

  if (!is.null(optimal_baseline_data)) {
    p = p + 
      annotate("segment",
               x=(1:5)-0.5, xend=(1:5)+0.5,
               y=optimal_baseline_data$expected_reward,
               yend=optimal_baseline_data$expected_reward,
               linetype=1, alpha=0.5) 
  }

  if (!is.null(other_baseline_data)) {
    p = p + 
      geom_spoke(data=other_baseline_data,
                 aes(x=game_type, y=expected_reward, angle=0, radius=1),
                 linetype=2, alpha=0.5,
                 position=position_nudge(x=-0.5)) 
  }

@@

```{r}
check_if_new_game = function(run_type, game_str) {
  case_when(grepl("su_heldout", run_type) ~ game_str %in% new_game_strs_su_heldout,
            grepl("only_one$", run_type) ~ game_str %in% new_game_strs_only_one,
            grepl("only_one_ho", run_type) ~ game_str %in% new_game_strs_only_one_ho,
            grepl("random", run_type) ~ game_str %in% new_game_strs_random_heldout,
            T ~ game_str %in% new_game_strs)
}
```

```{r}
baseline_data = read_csv("baseline_data.csv") %>%
  mutate(game_type = str_extract(game, "high_card|match|straight_flush|pairs_and_high|sum_under"),
         losers = grepl("l_1", game),
         black_valuable = grepl("bv_1", game),
         suits_rule = grepl("sr_1", game))
```
```{r}
optimal_baselines = baseline_data %>% 
  filter(policy == "optimal") %>%
  group_by(game_type) %>%
  summarize(expected_reward = mean(expected_reward)) %>%
  ungroup()
```

And "baselines" which show the performance using strategy from most correlated prior game

```{r}
most_correlated_baselines = crossing(run_type=result_subdirs, baseline_data) %>%
  mutate(is_new_game = check_if_new_game(run_type, game)) %>%
  filter((!is_new_game & policy == "optimal") | (policy != "optimal" & !check_if_new_game(run_type, policy))) %>%
  group_by(run_type, game, is_new_game, game_type, losers, black_valuable, suits_rule) %>%
  summarize(optimal_expected_reward = max(expected_reward)) %>%
  group_by(run_type, game_type, is_new_game) %>%
  summarize(expected_reward = mean(optimal_expected_reward)) %>%
  mutate(is_new_game = factor(is_new_game, levels=c(F, T), labels=c("Trained", "Held out")),
         named_run_type = name_run_type(run_type))

@@

reward_d = reward_d %>%
  mutate(is_su_ho=grepl("su_heldout", run_type),
         is_only_one=grepl("only_one$", run_type),
         is_only_one_ho=grepl("only_one_ho", run_type),
         is_random_ho=grepl("random", run_type),
         is_default=!(is_su_ho | is_random_ho | is_only_one | is_only_one_ho)) %>%
  gather(game_str, reward, starts_with("game")) %>%
  mutate(game_type = str_extract(game_str, "high_card|match|straight_flush|pairs_and_high|sum_under"),
         losers = grepl("l_1", game_str),
         black_valuable = grepl("bv_1", game_str),
         suits_rule = grepl("sr_1", game_str),
         is_new_game = check_if_new_game(run_type, game_str),
         is_new_game = factor(is_new_game, labels=c("Trained", "Held out")))
```

@@

```{r}

# summary_plot(reward_d %>%
#                filter(epoch == 0),
#              "game_type", "reward", "is_new_game",
#              optimal_baselines) +
#   facet_wrap(. ~ named_run_type)
```

```{r}
summary_plot(reward_d %>%
               filter(epoch == 0,
                      run_type == "basic"),
             "game_type", "reward", "is_new_game",
             optimal_baselines,
             most_correlated_baselines %>%
               filter(run_type == "basic",
                      is_new_game == "Held out")) 
#ggsave("~/Documents/grad/res/psych/ARRR/figures/meta_RL/basic.png", width=5, height=3)
```

```{r}
summary_plot(reward_d %>%
               filter(epoch == 0),
             "game_type", "reward", "is_new_game",
             optimal_baselines,
             most_correlated_baselines %>%
               filter(is_new_game == "Held out")) +
  facet_wrap(. ~ named_run_type) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#ggsave("writing/figures/basic_meta_learning.png", width=5, height=3.5)
#ggsave("~/Documents/grad/res/psych/ARRR/figures/meta_RL/basic_random_holdout_half.png", width=5, height=3.5)
```

@@

  mutate(is_su_ho=grepl("su_heldout", run_type),
         is_only_one=grepl("only_one$", run_type),
         is_only_one_ho=grepl("only_one_ho", run_type),
         is_random_ho=grepl("random", run_type),
         is_default=!(is_su_ho | is_random_ho | is_only_one | is_only_one_ho)) %>%
  gather(game_str, reward, starts_with("game")) %>%
  mutate(game_type = str_extract(game_str, "high_card|match|straight_flush|pairs_and_high|sum_under"),
         losers = grepl("l_1", game_str),
         black_valuable = grepl("bv_1", game_str),
         suits_rule = grepl("sr_1", game_str),
         is_new_game = check_if_new_game(run_type, game_str))

@@

#ggsave("~/Documents/grad/res/psych/ARRR/figures/meta_RL/basic_sweeping.png", width=5, height=3)
```

@@

```{r}
language_d = load_d(results_dir, c("language_noncompositional_old", 
                                   #"language",  "language_nontoggling_random_holdout", 
                                   "language_simplified", "language_meta_nontoggling", "language_noncompositional",
                                   "language_simplified_random_holdout"), num_runs, "new_language_rewards")

language_d = language_d %>%
  mutate(is_su_ho=grepl("su_heldout", run_type),
         is_only_one=grepl("only_one$", run_type),
         is_only_one_ho=grepl("only_one_ho", run_type),
         is_random_ho=grepl("random", run_type),
         is_default=!(is_su_ho | is_random_ho | is_only_one | is_only_one_ho)) %>%
  gather(game_str, reward, starts_with("game")) %>%
  mutate(game_type = str_extract(game_str, "high_card|match|straight_flush|pairs_and_high|sum_under"),
         losers = grepl("l_1", game_str),
         black_valuable = grepl("bv_1", game_str),
         suits_rule = grepl("sr_1", game_str),
         is_new_game = check_if_new_game(run_type, game_str))
```

```{r}
ggplot(language_d %>%
         filter(epoch == 0,
                is_new_game),
       aes(x=run, y=reward, fill=as.factor(run))) +
  geom_hline(yintercept=0.,
             linetype=2,
             alpha=0.5) +
  geom_violin(position="dodge") +
  geom_point(aes(group=as.factor(run)),
             size=3,
             stat="summary", fun.y="mean",
             position=position_dodge(width=0.9)) +
  scale_color_brewer(palette="Set2") + 
  ylim(-0.6, 0.6) +
  labs(x="Run",
       y="Reward density on held-out tasks") + 
  guides(fill=guide_legend(title="Run")) +
  facet_wrap(~named_run_type)
#ggsave("writing/figures/language_basic_by_run.png", width=5, height=3.5)
```

```{r}
summary_plot(language_d %>%
               filter(epoch == 0),
             "game_type", "reward", "is_new_game",
             optimal_baselines) +
  facet_wrap(. ~ named_run_type) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("writing/figures/language_basic.png", width=5, height=3.5)
#ggsave("~/Documents/grad/res/psych/ARRR/figures/meta_RL/crappy_language.png", width=5, height=3)
```

@@

```{r}
meta_d = load_d(results_dir, result_subdirs, num_runs, "new_meta_true_losses")

meta_d = meta_d %>%
  filter(epoch == 0) %>%
  mutate(is_su_ho=grepl("su_heldout", run_type),
         is_only_one=grepl("only_one$", run_type),
         is_only_one_ho=grepl("only_one_ho", run_type),
         is_random_ho=grepl("random", run_type),
         is_default=!(is_su_ho | is_random_ho | is_only_one)) %>%
  gather(map_str, reward, -c(run, run_type, named_run_type, epoch), -starts_with("is_")) %>%
  separate(map_str, c("meta_task", "source", "target"), sep="[\\.]+") %>%
  mutate(is_new = check_if_new_game(run_type, source) | check_if_new_game(run_type, target),
         is_new = factor(is_new, labels=c("Trained", "Held out")))
```

@@

#ggsave("~/Documents/grad/res/psych/ARRR/figures/meta_RL/basic_basic_random_meta.png", width=5, height=3.5)
#ggsave("writing/figures/meta_mapping.png", width=5, height=3.5)
```

@@

```{r}
lang_meta_d = load_d(results_dir, c(
                              # "language", "language_simplified", "language_meta_nontoggling", "language_nontoggling_random_holdout", 
                               "language_noncompositional_old", "language_simplified", "language_simplified_random_holdout",
                               "language_meta_nontoggling", "language_noncompositional"
                               ), num_runs, "new_language_meta_true_losses")

lang_meta_d = lang_meta_d %>%
  filter(epoch == 0) %>%
  mutate(is_su_ho=grepl("su_heldout", run_type),
         is_only_one=grepl("only_one$", run_type),
         is_only_one_ho=grepl("only_one_ho", run_type),
         is_random_ho=grepl("random", run_type),
         is_default=!(is_su_ho | is_random_ho | is_only_one)) %>%
  gather(map_str, reward, -c(run, run_type, named_run_type, epoch), -starts_with("is_")) %>%
  separate(map_str, c("meta_task", "source", "target"), sep="[\\.]+") %>%
  mutate(is_new = check_if_new_game(run_type, source) | check_if_new_game(run_type, target),
         is_new = factor(is_new, labels=c("Trained", "Held out")))
```

```{r}
summary_plot(lang_meta_d %>%
               filter(epoch == 0),
             "meta_task", "reward", "is_new") +
  facet_wrap(. ~ named_run_type)  +
  guides(color=guide_legend(title=NULL)) +
  labs(x="Game type", y="Average reward") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 #ggsave("writing/figures/language_meta_mapping.png", width=5, height=3.5)
```
# embeddings
